\documentclass[11pt,english]{article}
\usepackage[a4paper, total={6in, 8in}]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{babel}
\usepackage{xcolor}
\usepackage{url}
\usepackage{csquotes}
\usepackage{pgfplots}
\usepackage[style=numeric-comp]{biblatex}
\addbibresource{Resources/bibliography.bib}


%%% Custom definitions %%%
% Shorthands
\newcommand{\ie}{i.\@e.~}
\newcommand{\eg}{e.\@g.~}
\newcommand{\wrt}{w.\@r.\@t.~}
\newcommand{\idea}[1]{\color{blue}{#1}\color{black}}
\newcommand{\temp}[1]{\color{green}{#1}\color{black}}
\newcommand{\chapref}[1]{Chapter \ref{#1}}

\newcommand{\figref}[1]{\figurename~\ref{#1}}


\title{Ethical trade-offs in medical Machine Learning technologies}
\author{
    Albrecht, Thomas (5733587)
    \and
    Petruck, Julian (3857386)
    \and
    Jaques, Arthur (5998179)}



\begin{document}

\pagenumbering{gobble}

\maketitle

\begin{abstract}
We use the application of Machine Learning (ML) to healthcare as a case study of ethical trade-offs.
We first introduce the concept of Pareto optimality to formalize trade-offs in general.
Then we identify three main trade-offs in fair ML: The choice of fairness measure and the trade-offs between accuracy, fairness and privacy. 
We also discuss some ideas that have been introduced to handle them.
We examine the particular trade-offs that appear in medical practice, be it in policy, treatment, or research. Because there is psychological resistance to trade-off talk in health care since it touches on the `sacred value' of life, such trade-offs are often identified as moral dilemmas. We show that fairness problems have been identified in unaided medicine, and many issues identified in the fair ML literature are just their derivations. After arguing for the need of ethical assessments of ML tools that are relative to clinicians, we analyze what differences ML decision support tools have to human doctors with respect to fairness, and what they can positively contribute.
Finally, we look beyond the issue of fairness towards other ethical trade-offs. Our analysis shows that while very often trade-offs at the intersection of the fields of medicine and ML look like new problems, most of them can be related to trade-offs we are used to dealing with in other technologies.
\end{abstract}


\clearpage
\setcounter{tocdepth}{2}
\tableofcontents
\clearpage


\pagenumbering{arabic}
\section{Introduction}
\label{introduction}
\input{chapters/introduction}

\section{Introduction to trade-offs}
\label{tradeoff}
\input{chapters/tradeoff.tex}

\section{Trade-offs in medicine}
\label{medicine}
\input{chapters/medicine}

\section{Fairness in medical Machine Learning systems}
\label{combination}
\input{chapters/combination}

\section{Further ethical trade-offs in medical Machine Learning systems}
\label{new_tradeoff}
\input{chapters/new_tradeoff.tex}

\section{Conclusions}
\label{conclusions}
\input{chapters/conclusion}

%\section{Arthur's ideas}
%\input{chapters/arthur_ideas.tex}
\clearpage


\pagenumbering{gobble}


\printbibliography

\end{document}